SylvanSeeds, a seed germination database for temperate deciduous forests

### 2020-09-15 README.Rmd
3rd submission to JVS 🌳🍋

![SylvanSeeds](data/2017070313 Blue Ridge Parkway small.jpg)

@@

This repository stores all the information related to the publication *SylvanSeeds, a seed germination database for temperate deciduous forests*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

*SylvanSeeds* is a dataset of seed germination ecology for the temperate broadleaf and mixed forest biome of the northern hemisphere. It contains data on >300 plant species, obtained from >500 published references. The data is in the form of germination proportions recorded under experimental combinations of temperature, light, scarification and stratification.

@@

Please cite the dataset as

>>>>

![SylvanSeeds](data/2017070313 Blue Ridge Parkway small.JPG)

@@

[*SylvanSeeds*](https://efernandezpascual.github.io/home/sylvanseeds.html) is a database of seed germination ecology for the temperate broadleaf and mixed forest biome of the northern hemisphere. It contains data on >300 plant species, obtained from >500 published references. The data is in the form of germination proportions recorded under experimental combinations of temperature, light, scarification and stratification.

This repository stores all the information related to the publication *SylvanSeeds, a seed germination database for temperate deciduous forests*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript. It also stores the code of the *SylvanSeeds* `shiny` app, that can be used to visualize the database.

@@

Please cite `SylvanSeeds` as:

### 2020-09-15 doc/manuscript.Rmd
3rd submission to JVS 🌳🍋

Seed traits have functional significance in all levels of plant biology, but there is a lack of germination databases of wide geographical scope. This article presents *SylvanSeeds*, a first global database of germination records for an ecologically coherent unit: Temperate Broadleaf and Mixed Forests. Data were gathered with a systematic literature search. A list of frequent taxa of the study area was created using `r MSplots` vegetation relevés from the sPlot database. The list was searched in the Web of Science. 6,791 references were screened, finding 555 articles from which data were extracted. *SylvanSeeds* includes `r MSrecords` germination records of `r MSspecies` species from `r MSfamilies` families (gymnosperms and angiosperms), collected in 46 countries between 1920-2017. It provides raw data for meta-analysis: proportions of seeds germinated in laboratory experiments of scarification, stratification, light-darkness, and constant-alternating temperatures. *SylvanSeeds* is freely distributed as a *.csv*. The shiny app *SylvanSeeds* (https://efernandezpascual.shinyapps.io/sylvanseeds/) is also presented, to make data accessible to the public. *SylvanSeeds* advances functional seed ecology and brings two innovations to plant science.  First, the data-gathering methodology can be extended to other biomes. Second, database and app can be a standard in further efforts to compile germination data.

@@

The database provided in this article, and the web app to visualize it, can have wide applicability in science and beyond. It can extend current trends in global analyses of plant traits at species [@RN3318] and community levels [@RN4664] to seed germination. The ecological determinants of seed germination are also valuable information for species distribution models [@RN3078]. The visualization of the database with the *SylvanSeeds* app can help plant ecologists select experimental treatments that are adequate for their experiments [@RN2999]. Outside of academia, germination data are useful for seed industries [@RN4665], restoration practitioners [@RN2241] and the implementation of regional schemes for seed-based landscape intervention [@RN4666]. The app is accessible to citizens in general who are interested in germinating wild plants. Finally, apart from the database itself, this article can bring two innovations to the seed ecology community, helping to advance the agenda of functional seed ecology [@RN4655]. First, the meta-analysis-inspired methodology used to compile the database can be extended to other biomes and lists of species, contributing to the creation of a global database for ecologically and biogeographically coherent floras. Second, the database and the app can serve as a standard in further efforts to compile and standardize seed germination data. The version of the database presented here is a first building block that will be expanded by updating it with more records and integrating it with datasets from other biomes and regions. The website https://efernandezpascual.shinyapps.io/sylvanseeds/ will publish and describe updates. All people interested in contributing to the growth of this germination database are encouraged to contact the corresponding author.

@@

All persons can use the database providing they cite this paper properly in any publications or in the metadata of any derived products that are produced using the database. The database is provided as Appendix S5 of this manuscript, but please do visit https://efernandezpascual.shinyapps.io/sylvanseeds/ to check for updates. It can be visualized with the *SylvanSeeds* app at  https://efernandezpascual.shinyapps.io/sylvanseeds/. The code of the app is stored at (note: the GitHub repo is kept private until publication of the manuscript).

@@

```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Geographical distribution of the germination records in the database. Each golden circle is a record. The green areas correspond to the extension of the Temperate Broadleaf and Mixed Forests biome according to the *WWF*.", fig.align = "center", fig.width=7, fig.height=3}
p1
```

```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Example of the germination records as shown by the *SylvanSeeds* app. Records for one species, the European pedunculate oak, *Quercus robur*. Each panel shows the results for a combination of experimental conditions, with the germination temperature varying within each panel. Bars represent the mean germination proportion and brackets the 95% binomial confidence interval.", fig.align = "center", fig.width=7, fig.height=9}
p2
```

>>>>

Seed traits have functional significance in all levels of plant biology, but there is a lack of germination databases of wide geographical scope. This report presents *SylvanSeeds* (https://efernandezpascual.github.io/home/sylvanseeds.html), a first global database of germination records for an ecologically coherent unit: Temperate Broadleaf and Mixed Forests. Data were gathered with a systematic literature search. A list of frequent taxa of the study area was created using `r MSplots` vegetation relevés from the sPlot database. The list was searched in the Web of Science. 6,791 references were screened, finding 555 articles from which data were extracted. *SylvanSeeds* includes `r MSrecords` germination records of `r MSspecies` species from `r MSfamilies` families (gymnosperms and angiosperms), collected in 46 countries between 1920-2017. It provides raw data for meta-analysis: proportions of seeds germinated in laboratory experiments of scarification, stratification, light-darkness, and constant-alternating temperatures. *SylvanSeeds* is freely distributed as a *.csv*. A shiny web app is also presented, to make data accessible to the public. *SylvanSeeds* advances functional seed ecology and brings two innovations to plant science.  First, the data-gathering methodology can be extended to other biomes. Second, database and app can be a standard in further efforts to compile germination data.

@@

The database provided in this article, and the web app to visualize it, can have wide applicability in science and beyond. It can extend current trends in global analyses of plant traits at species [@RN3318] and community levels [@RN4664] to seed germination. The ecological determinants of seed germination are also valuable information for species distribution models [@RN3078]. The visualization of the database with the *SylvanSeeds* app can help plant ecologists select experimental treatments that are adequate for their experiments [@RN2999]. Outside of academia, germination data are useful for seed industries [@RN4665], restoration practitioners [@RN2241] and the implementation of regional schemes for seed-based landscape intervention [@RN4666]. The app is accessible to citizens in general who are interested in germinating wild plants. Finally, apart from the database itself, this article can bring two innovations to the seed ecology community, helping to advance the agenda of functional seed ecology [@RN4655]. First, the meta-analysis-inspired methodology used to compile the database can be extended to other biomes and lists of species, contributing to the creation of a global database for ecologically and biogeographically coherent floras. Second, the database and the app can serve as a standard in further efforts to compile and standardize seed germination data. The version of the database presented here is a first building block that will be expanded by updating it with more records and integrating it with datasets from other biomes and regions. The website https://efernandezpascual.github.io/home/sylvanseeds.html will publish and describe updates. All people interested in contributing to the growth of this germination database are encouraged to contact the corresponding author.

@@

All persons can use the database providing they cite this paper properly in any publications or in the metadata of any derived products that are produced using the database. The database is provided as **Appendix S5** of this manuscript, but please do visit https://efernandezpascual.github.io/home/sylvanseeds.html to check for updates. It can be visualized with the *SylvanSeeds* app at  https://efernandezpascual.shinyapps.io/sylvanseeds/. The code to create this manuscript and the app is stored at https://github.com/efernandezpascual/sylvanseeds.

@@

```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Geographical distribution of the germination records in the database. Each golden circle is a record. The green areas correspond to the extension of the Temperate Broadleaf and Mixed Forests biome according to the *WWF*.", fig.width=7, fig.height=3}
p1
```

```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Example of the germination records as shown by the *SylvanSeeds* app. Records for one species, the European common beech, *Fagus sylvatica*. Each panel shows the results for a combination of experimental conditions, with the germination temperature varying within each panel. Bars represent the mean germination proportion and brackets the 95% binomial confidence interval.", fig.width=7, fig.height=5}
p2
```
